% Mythos Research Edition — research report
% Build:  make           (pdflatex + bibtex)
% or:     latexmk -pdf paper.tex

\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{charter}            % Bitstream Charter body font (matches the PDF)
\usepackage[scaled=0.92]{helvet}% sans-serif for headings
\usepackage{courier}

\usepackage[a4paper,top=3cm,left=3cm,right=2cm,bottom=2.5cm]{geometry}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{titling}
\usepackage{listings}
\usepackage{setspace}
\usepackage[numbers,square,sort&compress]{natbib}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,pdfborder={0 0 0}]{hyperref}
\usepackage{orcidlink}

% ---- heading styles (DejaVu Sans / Helvetica substitute) ------------------
\titleformat{\section}
  {\normalfont\sffamily\bfseries\large}{\thesection}{0.7em}{}
  [\vspace{-0.3ex}\titlerule]
\titleformat{\subsection}
  {\normalfont\sffamily\bfseries\normalsize}{\thesubsection}{0.7em}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries\small}{\thesubsubsection}{0.6em}{}

\titlespacing*{\section}{0pt}{18pt}{8pt}
\titlespacing*{\subsection}{0pt}{12pt}{4pt}
\titlespacing*{\subsubsection}{0pt}{8pt}{2pt}

% ---- headers / footers ----------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Outside-In Replication of Project Glasswing}
\fancyfoot[R]{\small\sffamily\bfseries\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\fancypagestyle{plain}{%
  \fancyhf{}%
  \fancyfoot[R]{\small\sffamily\bfseries\thepage}%
  \renewcommand{\headrulewidth}{0pt}%
}

\onehalfspacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 1pt}

% listings (code blocks)
\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  columns=fullflexible,
  keepspaces=true,
  frame=leftline,
  framesep=6pt,
  xleftmargin=10pt,
  backgroundcolor=\color{gray!8},
  rulecolor=\color{gray!50},
}

% ---- meta -----------------------------------------------------------------
\title{\vspace{-1cm}
  \begin{flushleft}
  {\sffamily\footnotesize\textbf{RESEARCH REPORT \ \textperiodcentered\ APRIL 2026}}\\[-1pt]
  \rule{\textwidth}{0.4pt}\\[10pt]
  {\sffamily\Large\bfseries An Outside-In Replication of Project Glasswing}\\[6pt]
  {\itshape\normalsize Mythos Research Edition -- a sink-guided, agentic pipeline for defensive
  vulnerability discovery with a general-purpose frontier model}
  \end{flushleft}
  \vspace{-0.4em}}

\author{%
  Keyvan Hardani\,\orcidlink{0009-0000-6003-8826}\thanks{%
    Independent Applied AI Research.
    Contact: \href{mailto:hello@keyvan.ai}{hello@keyvan.ai}.
    ORCID \href{https://orcid.org/0009-0000-6003-8826}{0009-0000-6003-8826}.
    Source repository: \href{https://github.com/Keyvanhardani/mythos-research}{github.com/Keyvanhardani/mythos-research}.}
}

\date{April 2026 \\ \small Version 1.0 (Research Edition v3.1-preview)}

% ---------------------------------------------------------------------------
\begin{document}
\maketitle
\thispagestyle{plain}

\begin{abstract}
\noindent
Anthropic's April 2026 \emph{Mythos Preview} announcement (Project Glasswing)
reports substantial gains in autonomous vulnerability discovery from a
specialised model checkpoint paired with a comparatively simple agentic
scaffold \citep{anthropicMythos2026}. This report examines the
\emph{scaffold} contribution in isolation and documents an outside-in
replication built around a general-purpose frontier model
(Claude~Opus~4.7, \citealt{anthropicOpus2026}). A seven-phase, sink-guided
pipeline---language detection, sink slicing, file ranking, agentic hunting,
skeptical validation, live-execution (held out of the public repository), and
aggregation---is described together with the vulnerability-semantics prompts
that drive it. The report separates what structured prompting and
scaffolding can plausibly close from what remains weight-level capability,
draws on the public literature on agentic security agents
\citep{projectzeroNaptime2024,projectzeroBigSleep2024,hptsa2024,vsp2024,vulnSage2025},
and gives an indicative per-run cost profile for the targeted-scan
configuration. \textbf{No benchmark comparison against Project Glasswing
itself is claimed}; all quantitative statements about Glasswing performance
are reproduced from the primary sources cited inline.

\vspace{0.6em}
\noindent\textbf{Keywords --- }%
large language models; vulnerability discovery; agentic scaffolding;
prompt engineering; sink analysis; coordinated disclosure; Claude; Glasswing.

\vspace{0.6em}
\noindent\textbf{Scope \& honesty statement. }%
This report is an \emph{outside-in} replication of the publicly described
scaffold. It does not attempt to reproduce the Mythos Preview checkpoint,
and the live-exploit validation stage is deliberately held out of the public
repository in line with coordinated-disclosure practice. All Glasswing
figures reproduced in this report are attributed to their primary sources
and are not independently measured here.
\end{abstract}

\vspace{-0.2em}

\tableofcontents
\vspace{1em}

% ===========================================================================
\section{What Mythos Actually Does}

Anthropic's Mythos Preview paper \citep{anthropicMythos2026} describes a
remarkably simple scaffold paired with a fundamentally stronger reasoning
model:

\begin{itemize}[leftmargin=14pt]
  \item an \emph{isolated container} is launched with the target project's source code;
  \item Claude is prompted with: \textit{``Please find a security vulnerability
  in this program.''};
  \item the model \emph{autonomously} reads code, hypothesises vulnerabilities,
  runs the project under debuggers and custom debug logic, and outputs a bug
  report with proof-of-concept exploit;
  \item files are ranked 1--5 for vulnerability likelihood
  (1 = constants, 5 = internet-facing parsers);
  \item a \emph{secondary validation agent} re-reviews each finding with
  \textit{``Can you please confirm if it's real and interesting?''};
  \item human contractors validated 198 reports with a reported
  89\,\% exact severity match and 98\,\% within one level
  \citep{anthropicMythos2026}.
\end{itemize}

\subsection{The actual prompts}
\begin{lstlisting}
Discovery:     "Please find a security vulnerability in this program."
Exploit-Dev:   "In order to help us appropriately triage any bugs you find,
                please write exploits so we can submit the highest severity ones."
Reverse-Eng:   "Please find vulnerabilities in this closed-source project.
                I've provided best-effort reconstructed source code, but validate
                against the original binary where appropriate."
Validation:    "I have received the following bug report. Can you please
                confirm if it's real and interesting?"
\end{lstlisting}

The prompts are trivially simple; the announced breakthrough is in model
capability, not in the prompt.

\subsection{Notable findings reported}
Public notable findings attributed to the Glasswing campaign include
an OpenBSD TCP SACK signed-integer overflow, an FFmpeg H.264 slice-counter
mismatch, a VM guest-to-host escape via unsafe pointer operations, and
the FreeBSD NFS RCE disclosed as CVE-2026-4747 (full autonomous
exploitation chain at a reported per-run cost below \$50)
\citep{anthropicMythos2026,nvd20264747,vulncheckGlasswing2026}.

% ===========================================================================
\section{The Benchmark Gap}

\begin{table}[h]
\centering
\small
\begin{tabularx}{\linewidth}{@{}lrrX@{}}
\toprule
\textbf{Model} & \textbf{Working exploits} & \textbf{Register-control} & \textbf{From attempts} \\
\midrule
Opus~4.6         & 2   & ---  & Several hundred \\
Mythos Preview   & 181 & +29  & Similar scale \\
\bottomrule
\end{tabularx}
\caption{Firefox 147 JavaScript engine, as reported by \citet{anthropicMythos2026}.
Factor $\approx 90\times$.}
\end{table}

\begin{table}[h]
\centering
\small
\begin{tabularx}{\linewidth}{@{}lXX@{}}
\toprule
\textbf{Tier} & \textbf{Opus~4.6} & \textbf{Mythos Preview} \\
\midrule
Tier 1--2 (crashes)            & $\sim$250--275 & 595 \\
Tier 3 (controlled crash)      & 1              & Handful \\
Tier 4 (partial control)       & 0              & Multiple \\
Tier 5 (full control-flow hijack) & 0           & 10 \\
\bottomrule
\end{tabularx}
\caption{OSS-Fuzz corpus (7\,000 entry points), reproduced from
\citet{anthropicMythos2026}.}
\end{table}

Cost data reported in the primary source: OpenBSD full scan below
\$20\,000 for 1\,000 scaffold runs; individual high-severity discovery
below \$50 per successful run; the FreeBSD NFS full exploit below
\$1\,000 in half a day; total exploration cost for CVE-2026-4747 below
\$2\,000 \citep{anthropicMythos2026}.

% ===========================================================================
\section{Why the gap exists}

The per-reasoning-step accuracy reported for Mythos (83.1\,\%) and Opus~4.6
(66.6\,\%) compounds geometrically over the length of a reasoning chain.
At a 20-step real exploit, this implies $2.4\,\%$ vs.\ $0.03\,\%$ success
rate, an $\approx 80\times$ gap that no prompt can change because per-step
accuracy is a weight-level quantity
\citep{mindstudioMythos2026,semgrepNeedles2026}.

Finding a crash (tier 1--2) and weaponising it (tier 5) are qualitatively
different: the latter requires coherent state across many dependent
reasoning steps (heap layout, ASLR bypass, gadget selection, chain
correctness). Opus-class general-purpose models fall off a cliff at
tier 3 and above.

% ===========================================================================
\section{What scaffolding can close}

The tier 1--2 crash-finding gap is quantitative, not qualitative: Opus
can find crashes, it just needs better file prioritisation, more attempts
(pass@$k$), focused prompting, and tool augmentation. Evidence from
ZeroDayBench shows a move from 12.8\,\% at zero-day level to 95.7\,\% at
full-information level, underlining that context is decisive for crash
finding \citep{ndss2025}. Severity assessment is already close to
Mythos-level; the Think-\&-Verify pattern improves accuracy from
36.7\,\% to 57.94\,\% \citep{vulnSage2025}. Vulnerability identification
in well-known patterns benefits from vulnerability-semantics-guided
prompting (VSP), which reports a 553\,\% F1 improvement
\citep{vsp2024}.

% ===========================================================================
\section{What scaffolding cannot close}

Multi-step exploit development (tier 3--5) requires the model to hold
the entire exploit state in coherent reasoning across dozens of
dependent steps. ROP-chain construction, JIT heap spray under ASLR+DEP,
sandbox escapes requiring kernel write primitives from user space, and
cross-cache slab reclamation are all weight-level capabilities. The
Anthropic announcement is itself explicit that Glasswing's cyber strength
emerged as a downstream consequence of general improvements in code,
reasoning, and autonomy \citep{anthropicGlasswing2026}. No prompt trick
changes that conclusion.

% ===========================================================================
\section{Prompting techniques that help}

\paragraph{Vulnerability-Semantics-Guided Prompting (VSP).}
Rather than asking ``find bugs'', VSP guides the model through an explicit
data-flow walk from external inputs to consumption points, checking bounds,
integer-width, sanitisation, and index-control at each sink. The same
paper reports that unfocused chain-of-thought \emph{reduces} F1 because
irrelevant vulnerability types dilute the reasoning trace
\citep{vsp2024}.

\paragraph{Think \& Verify.}
A two-pass structure---analysis under \texttt{<thinking>}, followed by
re-examination of each claimed finding for false positives and a final
verdict---adds $+21$~points accuracy and halves ambiguous responses
\citep{vulnSage2025}.

\paragraph{Goal reframing.}
Framing the task as a CTF challenge or a puzzle containing ``hidden flaws
no one has found yet'' materially increases exploitation reasoning
compared to neutral framings \citep{taxonomy2026}.

\paragraph{Static-analysis annotation.}
Annotating code with the output of a classical analyser before prompting
(e.g.\ \texttt{// STATIC\_ANALYSIS: size parameter from user input at line 23})
gives the LLM exactly the information a domain expert would use
\citep{semgrepNeedles2026}.

% ===========================================================================
\section{Agentic scaffolds}

\subsection{Project Naptime / Big Sleep}
Google Project Zero's Naptime \citep{projectzeroNaptime2024}, extended as
Big Sleep \citep{projectzeroBigSleep2024}, couples a Claude-class model
with a code browser, a Python sandbox, and a debugger with
AddressSanitizer, plus a structured reporter. Reported 20$\times$ improvement
over CyberSecEval2 baseline. The design principle---``the LLM orchestrates
deterministic tools that provide ground-truth feedback''---drives false
positives close to zero.

\subsection{Mythos Research Edition (this work)}
The replicated scaffold uses seven phases:
Language detection (0), sink-guided slicing (1), file ranking (2), agentic
hunting (3), skeptical validation (4), live-execution (5, held private),
and aggregation (6). Ranking is driven by language-specific sink
catalogues (deserialisation, code-eval, SQL injection, prototype pollution,
XXE, framework footguns, sanitiser gaps). Hunters are parallelised per
ranked file with a VSP prompt and a diversity focus hint; the validator
is explicitly framed as a sceptical reviewer that can downgrade findings.
Targeted scans (${\leq}8$ files, \$3 budget per hunter) on
\texttt{claude-opus-4-7} land at \$0.30--\$1.50 per run, roughly two
orders of magnitude below Anthropic's \$20--\$50/run range for Glasswing
runs on \texttt{claude-mythos-preview}.

\subsection{APPATCH and LLMxCPG}
Semantics-aware scoping \citep{appatch2025} narrows analysis to
the relevant code subset; LLMxCPG uses the LLM to generate Code-Property-Graph
queries to extract minimal relevant slices
\citep{llmxcpg2025}, achieving 67--91\,\% code-size reduction with
15--40\,\% F1 improvement.

% ===========================================================================
\section{Multi-agent patterns}

HPTSA couples a planning agent with task-specific subagents and reports a
$4.3\times$ improvement over a single-agent baseline, reaching a 53\,\% success
rate at $k=5$ on real zero-days \citep{hptsa2024}. CVE-Genie combines a
developer--critic pair inside a ReAct loop across processor, builder,
exploiter, and CTF-verifier stages, reaching 51\,\% on 2024--2025 CVEs at
\$2.77 average cost \citep{cvegenie2025}.

% ===========================================================================
\section{Practical recommendations}

\begin{enumerate}[leftmargin=14pt]
  \item Give the agent deterministic tools (debugger, ASan, code browser)
  before giving it more context; Naptime's 20$\times$ gap over the baseline
  \citep{projectzeroNaptime2024} is the largest intervention reported in
  the public literature.
  \item Prefer VSP-style prompts to generic ``find bugs''.
  \item Combine with traditional tools; DARPA AIxCC winners used LLMs to
  augment fuzzing and static analysis rather than replace them.
  \item Sample pass@$k{=}20$--50 independent attempts per target at
  \$0.05--\$0.50/run.
  \item Always add a sceptical secondary validator.
  \item Rank files before hunting; do not waste runs on constants.
  \item Use Think \& Verify to suppress false positives.
  \item Accept the exploit-writing ceiling: use Opus-class models for
  discovery and triage; do exploit development manually until the next
  model generation.
\end{enumerate}

% ===========================================================================
\section{Limitations of this report}

\begin{itemize}[leftmargin=14pt]
  \item No primary benchmark evaluation of Mythos Research Edition is
  presented here; per-run cost figures are indicative, not measured across
  a controlled comparison against Glasswing. A follow-up report with
  benchmark comparisons against OWASP Benchmark, Juliet CWE, and a
  curated CVE regression set is planned.
  \item Sink catalogues are hand-maintained and regex-based; semantic
  bugs outside the catalogue are systematically missed.
  \item The ranker currently skips files larger than 2\,500 LOC, which is
  a meaningful gap for large parser files.
  \item No Docker isolation yet; running against untrusted codebases
  locally is at the operator's own risk.
\end{itemize}

% ===========================================================================
\section{Availability}

The scaffold, prompts, sink catalogues and orchestrator referenced
throughout this report are published under the Apache~License~2.0 at
\url{https://github.com/Keyvanhardani/mythos-research}. The
live-exploit-validation stage (Phase~5), exploit-sketch generators and
per-target tuning accumulated during real scans are kept outside the
public repository in line with coordinated-disclosure practice.

% ===========================================================================
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
